---
title: 'Briefing for C19AG: Comparison of doubling times'
subtitle: 'NOT FOR DISTRIBUTION'
author: "COVID-19 Epidemic Response Unit, University of Edinburgh"
date: "02/04/2020"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

setwd('/Users/s1687811/Documents/GitHub/covid19/AG_briefing/')
today<- Sys.Date() - 2

# SOURCED SCRIPTS & ENVIRONMENTS
# 1) load packages + used functions
source('/Users/s1687811/Documents/GitHub/covid19/script/sourced_functions_doublingTime_reports.R') 
# 2) output of analysis script --> all content to run plot commands are in there --> May be used to re-work figures in a classic R script

load(paste0('/Users/s1687811/Documents/GitHub/covid19/AG_briefing/output/AG_briefing_analysis_output_', today, '.RData'))

# GENERAL FIGURES PARAMETERS
xlabels.font = 2
ylabels.font = 2
xlabels.cex = 1.3
ylabels.cex = 1.3

ylabels.line = 2.2
xlabels.line = 2.2

xticks.cex = 1.3
yticks.cex = 1.3

xticks.font = 2
yticks.font = 2

```


# Key points summary

We compare the size and rate of increase of the COVID-19 epidemic for Scotland, London and the rest of the UK except for London (rUKxL).  
&nbsp;
 
**The epidemic in Scotland is ~`r abs(round(Scotland_vs_London, 0))` days `r text.Scotland_vs_London` London and is now growing at a faster rate.**  

&nbsp;
 
Based on deaths:  
&nbsp;
 
* The current doubling time for deaths in Scotland is `r Td.report[Td.report$variable == 'Scotland death','Td.obs']` days (95% confidence interval: `r Td.report[Td.report$variable == 'Scotland death','ci.low']` - `r Td.report[Td.report$variable == 'Scotland death','ci.upp']` days) (Figure 1).  
* This is not significantly different from doubling time for previous 7 days (2.4 days; 95%CI: 1.6-3.7 days).  
&nbsp;
 
Based on case counts per 10,000 population available as of `r format(today, "%d/%m/%Y")`:  
&nbsp;
 
* The epidemic in Scotland is `r abs(Scotland_vs_London)` days `r text.Scotland_vs_London`  London and `r abs(Scotland_vs_rUKxL)` days `r text.Scotland_vs_rUKxL` of rUKxL (Figure 2).  
* The current 7-day doubling time in Scotland is `r Td.report[Td.report$variable == 'Scotland cases','Td.obs']` days (95%CI: `r Td.report[Td.report$variable == 'Scotland cases','ci.low']` - `r Td.report[Td.report$variable == 'Scotland cases','ci.upp']` days).  
* This is very similar to the doubling time for previous 7 days (4.2 days; 95%CI: 3.9-4.5 days).  
* The current doubling time in Scotland is significantly faster than London  (`r Td.report[Td.report$variable == 'London cases','Td.obs']` days, 95%CI: `r Td.report[Td.report$variable == 'London cases','ci.low']` - `r Td.report[Td.report$variable == 'London cases','ci.upp']` days) and significantly slower than rUKxL (`r Td.report[Td.report$variable == 'Rest of UK cases','Td.obs']` days, 95%CI: `r Td.report[Td.report$variable == 'Rest of UK cases','ci.low']` - `r Td.report[Td.report$variable == 'Rest of UK cases','ci.upp']` days) over the same time period.  
* Across Health Boards in Scotland there is variation in cumulative case incidence (`r round(min(tail(df.fig3b, 1)[,-1]), 1)` to `r round(max(tail(df.fig3b, 1)[,-1]), 1)` per 10,000 population, Figures 3, 4) and doubling time (`r min(Td.report$Td.obs)` to `r max(Td.report$Td.obs)` days, Figure 5).  
  
# Results

```{r fig1, fig.height = 6, fig.width = 6, fig.align = "center"}

# Set up figure parameters
yaxti.f1 = 10
xaxti.f1 = 7

yseq = seq(0, round_any(max(d.death.sim.2$cumNumCases), 10, f = ceiling), yaxti.f1)
xseq = seq.Date(from = min(d.death.sim.2$date), to = max(d.death.sim.2$date), by = xaxti.f1)


# Draw plot, add axes labels, add axes ticks
plot(cumNumCases ~ date, data = d.death.sim.2, type = 'l', main = '  ',
     xlab = '', ylab = '', xaxt = 'n', yaxt = 'n')

mtext('Cumulative number of deaths', side = 2, font = ylabels.font, line = ylabels.line, cex = ylabels.cex)
mtext('Date (dd/mm)', side = 1, font = xlabels.font, line = xlabels.line, cex = xlabels.cex)

axis(2, at = yseq, labels = yseq, font = yticks.font, cex.axis = yticks.cex)
axis(1, at = xseq, labels = format(xseq, "%d/%m"), font = xticks.font, cex.axis = xticks.cex)


# Add shaded polygon on area over which Dt are calculated
polygon(x = c(t1, t2, t2, t1),
        y = c(head(yseq, 1), head(yseq, 1), tail(yseq, 1), tail(yseq, 1)),
        border = NA,
        col = adjustcolor('lightgrey', alpha = .4))


# Add background grids
abline(h = yseq, v = xseq, lty = 'dotted', col = 'lightgrey')

# Add legend
legend("topleft",
       legend = c("Doubling time (95%CI):",
                  paste('  ', Td.death.obs, ' (',ci.death.low, '-', ci.death.upp, ')')),
       col = 'black',
       ncol = 1,
       title.adj = 0.4,
       y.intersp = 1,
       text.font = c(2,1),
       text.width = 4,
       inset=c(-0.05,-0.005))
box()

# Replot line on top to make it more visible
lines(cumNumCases ~ date, data = d.death.sim.2, lwd = 2)

```
**Figure 1. Epidemic curve for Scotland based on deaths over time up to `r format(today, "%d/%m/%Y")`.** Doubling time estimated over the past 7 days is `r Td.report[1,'Td.obs']` days (95%CI: `Td.report[1,'ci.low']`-`Td.report[1,'ci.upp']` days).

&nbsp;


```{r fig2, fig.height = 6, fig.width = 12, fig.fullwidth = TRUE}

par(mfrow = c(1,2))

# FIG 2A ----
df.fig2A<-
  d.uk %>%
  rename(rUKxL = `Rest of UK`)


# Set up figure parameters
yaxti.f2 = 2000
xaxti.f2 = 7

yseq = seq(0, round_any(max(df.fig2A[,2:4]), 10, f = ceiling), yaxti.f2)
xseq = seq.Date(from = min(df.fig2A$date), to = max(df.fig2A$date), by = xaxti.f2)

# Draw plot, axes labels, axes ticks, background grids
plot(rUKxL ~ date, data = df.fig2A, type = 'l', main = '  ',
     xlab = '', ylab = '', xaxt = 'n', yaxt = 'n', lwd = 2,
     xlim = range(xseq),
     ylim = range(yseq))

mtext('Cumulative number of cases', side = 2, font = ylabels.font, line = ylabels.line, cex = ylabels.cex)
mtext('Date (dd/mm)', side = 1, font = xlabels.font, line = xlabels.line, cex = xlabels.cex)

axis(2, at = yseq, labels = yseq, font = yticks.font, cex.axis = yticks.cex)
axis(1, at = xseq, labels = format(xseq, "%d/%m"), font = xticks.font, cex.axis = xticks.cex)


abline(h = yseq, v = xseq, col = 'lightgrey', lty = 'dotted')

# Add lines
lines(rUKxL ~ date, data = df.fig2A, lwd = 3, col = 'black')
lines(London ~ date, data = df.fig2A, lwd = 3, col = 'red')
lines(Scotland ~ date, data = df.fig2A, lwd = 3, col = 'dodgerblue')



mtext('A', side = 3, font = 2, line = 0.2, adj = 0, cex = 1.5)

legend("topleft",
       legend = c('Region/Country:', ' rUKxL', ' London', ' Scotland'),
       ncol = 1,
       title.adj = 0.4,
       y.intersp = 1,
       text.font = c(2,1,1,1),
       text.col = c('black', 'black', 'red', 'dodgerblue'),
       text.width = 4,
       inset=c(-0.05,-0.005))
box()


# FIG 2B ----

df.fig2B<-
  d.uk %>%
  gather('region', 'cumcases', 2:4) %>%
  left_join(pops.uk, by = 'region') %>%
  mutate(cumcases_10k = cumcases * (10000/popsize)) %>%
  mutate(log10_cumcases_10k = log10(cumcases_10k)) %>%
  select(date, region, log10_cumcases_10k) %>%
  spread(region, log10_cumcases_10k) %>%
  rename(rUKxL = `Rest of UK`)

# Set up figure parameters
yaxti.f2b = 1
xaxti.f2b = 7

yseq = seq(round_any(min(df.fig2B[,2:4]), 1, f = floor),
           round_any(max(df.fig2B[,2:4]), 1, f = ceiling),
           yaxti.f2b)
xseq = seq.Date(from = min(df.fig2A$date), to = max(df.fig2A$date), by = xaxti.f2b)

# Draw plot, axes labels, axes ticks, background grids
plot(rUKxL ~ date, data = df.fig2B, type = 'l', main = '  ',
     xlab = '', ylab = '', xaxt = 'n', yaxt = 'n', lwd = 2,
     xlim = range(xseq),
     ylim = range(yseq))

mtext('Cumulative cases per 10k pop.', side = 2, font = ylabels.font, line = ylabels.line, cex = ylabels.cex)
mtext('Date (dd/mm)', side = 1, font = xlabels.font, line = xlabels.line, cex = xlabels.cex)

axis(2, at = yseq, labels = 10^(yseq), font = yticks.font, cex.axis = yticks.cex)
axis(1, at = xseq, labels = format(xseq, "%d/%m"), font = xticks.font, cex.axis = xticks.cex)

abline(h = yseq, v = xseq, col = 'lightgrey', lty = 'dotted')

# Add shaded polygon on area over which Dt are calculated
polygon(x = c(t1, t2, t2, t1),
        y = c(head(yseq, 1), head(yseq, 1), tail(yseq, 1), tail(yseq, 1)),
        border = NA,
        col = adjustcolor('lightgrey', alpha = .4))

# Add lines
lines(rUKxL ~ date, data = df.fig2B, lwd = 2, col = 'black')
lines(London ~ date, data = df.fig2B, lwd = 2, col = 'red')
lines(Scotland ~ date, data = df.fig2B, lwd = 2, col = 'dodgerblue')

mtext('B', side = 3, font = 2, line = 0.2, adj = 0, cex = 1.5)


makeInsetText<- function(variable.inTd.report, Td.report, TextVarName){
  
  td = Td.report[Td.report$variable == variable.inTd.report, c('Td.obs', 'ci.low', 'ci.upp')]
                
  insetText.td<- paste0(TextVarName, ': ', td$Td.obs, ' (', td$ci.low , '-', td$ci.upp,')')

  return(insetText.td)
}
Td.report$variable<- as.character(Td.report$variable)

legend.text<- c('Doubling time (95%CI):',
makeInsetText('Rest of UK cases', Td.report, 'rUKxL'),
makeInsetText('London cases', Td.report, 'London'),
makeInsetText('Scotland cases', Td.report, 'Scotland'))

legend("topleft",
       legend = legend.text,
       ncol = 1,
       #title.adj = 0.3,
       y.intersp = 1,
       text.font = c(2,1,1,1),
       cex = 0.9,
       text.col = c('black', 'black', 'red', 'dodgerblue'),
       inset=c(-0.05,-0.005))
box()

```

**Figure 2. Comparison of epidemic curves for Scotland, London and rUKxL up to `r format(today, "%d/%m/%Y")`**. **A)** Cumulative reported cases. **B)** Cumulative cases per 10,000 population on a log10 scale. Inset shows corresponding doubling times (in days) over the past 7 days (with 95% confidence intervals).
&nbsp;
 

&nbsp;
 
&nbsp;
 
&nbsp;
 
 
```{r fig3, fig.height = 6, fig.width = 12, fig.fullwidth = TRUE}

# Assemble legend inset text directly from Td.report
hb.names<- c(
makeInsetText('Grampian Shetland and Orkney cases', Td.report, 'Grampian, Shetland and Orkney'),
makeInsetText('Highland and Western Isles cases', Td.report, 'Highland and Western Isles'),
makeInsetText('Dumfries and Galloway cases', Td.report, 'Dumfries and Galloway'),
makeInsetText('Greater Glasgow and Clyde cases', Td.report, 'Greater Glasgow and Clyde'),
makeInsetText('Lanarkshire cases', Td.report, 'Lanarkshire'),
makeInsetText('Forth Valley cases', Td.report, 'Forth Valley'),
makeInsetText('Tayside cases', Td.report, 'Tayside'),
makeInsetText('Lothian cases', Td.report, 'Lothian'),
makeInsetText('Borders cases', Td.report, 'Borders'),
makeInsetText('Ayrshire cases', Td.report, 'Ayrshire'),
makeInsetText('Fife cases', Td.report, 'Fife'))


par(mfrow = c(1,2))


# FIG 3A ----

# Set up figure parameters
yaxti.f3 = 200
xaxti.f3 = 7
yseq = seq(0, round_any(max(d.cases[,2:ncol(d.cases)]), 10, f = ceiling), yaxti.f3)
xseq = seq.Date(from = min(d.cases$date), to = max(d.cases$date), by = xaxti.f3)

# Draw plot, axes labels, axes ticks, background grids
plot(Ayrshire ~ date, data = d.cases, type = 'l', main = '  ',
     xlab = '', ylab = '', xaxt = 'n', yaxt = 'n', lwd = 2,
     xlim = range(xseq),
     ylim = range(yseq))

mtext('Cumulative number of cases', side = 2, font = ylabels.font, line = ylabels.line, cex = ylabels.cex)
mtext('Date (dd/mm)', side = 1, font = xlabels.font, line = xlabels.line, cex = xlabels.cex)

axis(2, at = yseq, labels = yseq, font = yticks.font, cex.axis = yticks.cex)
axis(1, at = xseq, labels = format(xseq, "%d/%m"), font = xticks.font, cex.axis = xticks.cex)

abline(h = yseq, v = xseq, col = 'lightgrey', lty = 'dotted')

# Add lines
figb.lwd = 3
lines(Ayrshire ~ date, data = d.cases, lwd = figb.lwd, col = '#88CCEE')
lines(Borders ~ date, data = d.cases, lwd = figb.lwd, col = '#CC6677')
lines(`Dumfries and Galloway` ~ date, data = d.cases, lwd = figb.lwd, col = '#DDCC77')
lines(Fife ~ date, data = d.cases, lwd = figb.lwd, col = '#117733')
lines(`Forth Valley` ~ date, data = d.cases, lwd = figb.lwd, col = '#332288')
lines(`Grampian Shetland and Orkney` ~ date, data = d.cases, lwd = figb.lwd, col = '#AA4499')
lines(`Greater Glasgow and Clyde` ~ date, data = d.cases, lwd = figb.lwd, col = '#44AA99')
lines(`Highland and Western Isles` ~ date, data = d.cases, lwd = figb.lwd, col = '#999933')
lines(Lanarkshire ~ date, data = d.cases, lwd = figb.lwd, col = '#882255')
lines(Lothian ~ date, data = d.cases, lwd = figb.lwd, col = '#661100')
lines(Tayside ~ date, data = d.cases, lwd = figb.lwd, col = '#888888')


mtext('A', side = 3, font = 2, line = 0.2, adj = 0, cex = 1.5)


cols.f3a<- c("#AA4499","#999933", "#DDCC77","#44AA99", "#882255", "#332288", "#888888",
             "#661100", "#CC6677", "#88CCEE", "#117733")

hb.names.a<- do.call('rbind', strsplit(hb.names, ':'))[,1]

legend("topleft",
       legend = c('Health board:', hb.names.a),
       ncol = 1,
       title.adj = 0.4,
       y.intersp = 0.5,
       text.font = c(2,1,1,1),
       text.col = c('black',cols.f3a),
       text.width = 4,
       cex = 0.8,
       #lwd = 2,
       #lty = 1,
       inset=c(-0.07,-0.005))
box()


# FIG 3B ----

# Set up figure parameters
yaxti.f3b = 1
xaxti.f3b = 7

df.fig3b[,2:ncol(df.fig3b)]<- log10(df.fig3b[,2:ncol(df.fig3b)])

yseq = seq(round_any(min(df.fig3b[,2:ncol(df.fig3b)], na.rm = TRUE), 1, f = floor),
           round_any(max(df.fig3b[,2:ncol(df.fig3b)], na.rm = TRUE), 1, f = ceiling),
           yaxti.f3b)
xseq = seq.Date(from = min(df.fig3b$date), to = max(df.fig3b$date), by = xaxti.f3b)

# Draw plot, axes labels, axes ticks, background grids
plot(Ayrshire ~ date, data = df.fig3b, type = 'l', main = '  ',
     xlab = '', ylab = '', xaxt = 'n', yaxt = 'n', lwd = 2,
     xlim = range(xseq),
     ylim = range(yseq))

mtext('Cumulative cases per 10k pop.', side = 2, font = ylabels.font, line = ylabels.line, cex = ylabels.cex)
mtext('Date (dd/mm)', side = 1, font = xlabels.font, line = xlabels.line, cex = xlabels.cex)

axis(2, at = yseq, labels = 10^(yseq), font = yticks.font, cex.axis = yticks.cex)
axis(1, at = xseq, labels = format(xseq, "%d/%m"), font = xticks.font, cex.axis = xticks.cex)

abline(h = yseq, v = xseq, col = 'lightgrey', lty = 'dotted')

# Add shaded polygon on area over which Dt are calculated
polygon(x = c(t1, t2, t2, t1),
        y = c(head(yseq, 1), head(yseq, 1), tail(yseq, 1), tail(yseq, 1)),
        border = NA,
        col = adjustcolor('lightgrey', alpha = .4))

mtext('B', side = 3, font = 2, line = 0.2, adj = 0, cex = 1.5)

lines(Ayrshire ~ date, data = df.fig3b, lwd = figb.lwd, col = '#88CCEE')
lines(Borders ~ date, data = df.fig3b, lwd = figb.lwd, col = '#CC6677')
lines(`Dumfries and Galloway` ~ date, data = df.fig3b, lwd = figb.lwd, col = '#DDCC77')
lines(Fife ~ date, data = df.fig3b, lwd = figb.lwd, col = '#117733')
lines(`Forth Valley` ~ date, data = df.fig3b, lwd = figb.lwd, col = '#332288')
lines(`Grampian Shetland and Orkney` ~ date, data = df.fig3b, lwd = figb.lwd, col = '#AA4499')
lines(`Greater Glasgow and Clyde` ~ date, data = df.fig3b, lwd = figb.lwd, col = '#44AA99')
lines(`Highland and Western Isles` ~ date, data = df.fig3b, lwd = figb.lwd, col = '#999933')
lines(Lanarkshire ~ date, data = df.fig3b, lwd = figb.lwd, col = '#882255')
lines(Lothian ~ date, data = df.fig3b, lwd = figb.lwd, col = '#661100')
lines(Tayside ~ date, data = df.fig3b, lwd = figb.lwd, col = '#888888')


legend("topleft",
       legend = c('Health board:', hb.names),
       ncol = 1,
       title.adj = 0.4,
       y.intersp = 0.5,
       text.font = c(2,1,1,1),
       text.col = c('black',cols.f3a),
       text.width = 4,
       cex = 0.8,
       #lwd = 2,
       #lty = 1,
       inset=c(-0.07,-0.005))
box()


```
&nbsp;

**Figure 3. Comparison of epidemic curves for all Scottish Health Boards up to `r format(today, "%d/%m/%Y")`**. **A)** Cumulative reported cases. **B)** Cumulative cases per 10,000 population on log10 scale. Inset shows corresponding doubling times (in days) estimated over the past 7 days with 95% confidence intervals.

\newpage

```{r, fig.height = 7.5, fig.width = 10, fig.align = "center", warning = FALSE}

# Setting up a list to store last row of each cumulative incidence dataframes with their simulations
dat.hist.cumsum10k<- vector('list', length = length(regions))

for(i in 1:length(sims.store)){
 
# Fetch simulation output of a health board, add HB name to keep track, rename 'cumNumCases' (which is observed data), take last row (i.e. incidence at present)
latest<- sims.store[[i]] %>%
  mutate(region = names(sims.store[i])) %>%
  rename(cumNumCase_10k = cumNumCases) %>% 
  tail(., 1)

# Extract the simulations from this (The V columns), and compute CIs of this  bootstrap distribution of incidence
cumNumCases_10k.sim.last<- as.numeric(latest[,which(substr(colnames(latest), 1, 1) == 'V')])
ci.low.cumNumCases_10k<- round(quantile(cumNumCases_10k.sim.last, c(0.05), method = 6), 1)[[1]]
ci.upp.cumNumCases_10k<- round(quantile(cumNumCases_10k.sim.last, c(0.95), method = 6), 1)[[1]]

# Store results
dat.hist.cumsum10k[[i]]<-
latest %>%
  select(date, region, cumNumCase_10k) %>%
  mutate(ci.low = ci.low.cumNumCases_10k,
         ci.upp = ci.upp.cumNumCases_10k)
}

# Collapse list in dataframe, rename Grampian for the plot, order HB names as factors ordered by descending order of cumulative incidence
df.hist.cumsum10k<- do.call('rbind', dat.hist.cumsum10k)
df.hist.cumsum10k$region[which(df.hist.cumsum10k$region == "Grampian Shetland and Orkney")]<- "Grampian, Shetland and Orkney"
df.hist.cumsum10k$region<- factor(df.hist.cumsum10k$region, levels = df.hist.cumsum10k$region[rev(order(df.hist.cumsum10k$cumNumCase_10k))])


# Plot histogram - Simple ggplot
ggplot(df.hist.cumsum10k, aes(x = region, y = cumNumCase_10k))+
  geom_histogram(stat = 'identity', width = 0.8, color = "black", fill = 'gray45')+
  geom_errorbar(aes(ymin = ci.low, ymax = ci.upp), width = 0.20, lwd = 1)+
  xlab('Health board') + ylab('Cumulative reported cases per 10k pop.\n ')+
  theme_bw()+
  theme(legend.position="none",
        panel.border= element_blank(),
        axis.text.y = element_text(face="bold", colour="black", size=16),
        axis.text.x = element_text(colour="black", face="bold", size=14, angle = 45, vjust=1, hjust=1),
        axis.title.y = element_text(face="bold", colour="black", size=15),
        axis.title.x = element_text(face="bold", colour="black", size=15),
        axis.line.y = element_line(color="black", size = 0.5),
        axis.line.x = element_line(color="black", size = 0.5),
        plot.title = element_text(lineheight=.8, face="bold", hjust = 0.5),
        panel.grid.major = element_line(color = 'grey', linetype = 'dotted'))

```
&nbsp;

**Figure 4. Cumulative incidence for all Scottish Health Boards up to `r format(today, "%d/%m/%Y")`**. The error bars show the 95%CI of the cumulative incidence per 10,000 population reached at last time point over the bootstrapped simulated datasets with Poisson error structure.

\newpage

```{r, fig.height = 7, fig.width = 9, fig.align = "center", warning = FALSE}


# Reformat the Td report table to use as basic for plotting Dt in correct order
Td.report.analyses.clean<- 
  Td.report.analyses %>%
  rename(`Country/Health board` = variable,
         Dt = Td.obs) %>%
  select(`Country/Health board`, Dt, ci.low, ci.upp)

Td.report.analyses.clean.hb<- Td.report.analyses.clean[-c(1:4),] %>%
  arrange(-Dt)


# Plot Doubling time:
# Figure parameters settings
xaxt.f5 = 1

xseq.min<- floor(-range(Td.report.analyses.clean.hb[,2:4])[c(2)])
xseq.max<- ceiling(-range(Td.report.analyses.clean.hb[,2:4])[c(1)])
xseq<- seq(xseq.min, xseq.max, by = xaxt.f5)

# Set plot area, ad axes
par(mar = c(4, 15, 2, 1))

plot('', xlim = range(xseq), ylim = c(0, nrow(Td.report.analyses.clean.hb)+1),
     xaxt = 'n', yaxt = 'n', xlab = '', ylab = '')
axis(1, at = xseq, labels = abs(xseq), font = xticks.font, cex.axis = xticks.cex)

# Add axes labels and ticks
hb.names<- str_replace(as.character(Td.report.analyses.clean.hb$`Country/Health board`), ' cases', '')
hb.names[which(hb.names == "Grampian Shetland and Orkney")]<- "Grampian, Shetland and Orkney"
ytick<- seq(1, nrow(Td.report.analyses.clean.hb), 1) # Add healthboard names
axis(side=2, at=ytick, labels = FALSE, tick = TRUE)
text(par("usr")[1], ytick,  
     labels = hb.names, offset = 1, pos = 2, xpd = TRUE, font = yticks.font, cex = yticks.cex)

mtext('Doubling time (days)', side = 1, font = 2, line = 2.5)

# Add shaded polygons to highlight Dt windows
for(x in 1:(length(xseq)-1)){
xseq.focal<- xseq[c(x, length(xseq))]
polygon(x = rep(xseq.focal, each = 2),
        y = c(0, nrow(Td.report.analyses.clean.hb)+1, nrow(Td.report.analyses.clean.hb)+1, 0),
        border = 'NA',
        col = adjustcolor('firebrick', alpha = .1))
}

# Add the doubling times points and 95%CI
# Looping through the Td.report to fetch each value
for(c in 1:nrow(Td.report.analyses.clean.hb)){
  points(x = -Td.report.analyses.clean.hb$Dt[c], y = c, pch = 16, cex = 1.3) # Dt point
  lines(y = c(c,c), x = -Td.report.analyses.clean.hb[c,c('ci.low', 'ci.upp')], lwd = 3) # CI horizontal bar
  lines(y = c(c+0.1, c-0.1), x = -rep(Td.report.analyses.clean.hb$ci.low[c], 2), lwd = 3) # CI vertical bars
  lines(y = c(c+0.1, c-0.1), x = -rep(Td.report.analyses.clean.hb$ci.upp[c], 2), lwd = 3) # CI vertical bars
}

```
&nbsp;

**Figure 5. Doubling time of cases**. Doubling times are calculated over a 7 day period up to `r format(today, "%d/%m/%Y")` Error bars indicate 95%CI.

\newpage

# Data

- Case counts for Scotland and for Scottish HBs from https://www.gov.scot/coronavirus-covid-19/ (accessed 1200 `r format(today, "%d/%m/%Y")`).  
- Case counts for London and rUK except London from https://www.arcgis.com/apps/opsdashboard/index.html#/f94c3c90da5b4e9f9a0b19484dd4bb14 (accessed 2000 `r format(today, "%d/%m/%Y")`).  
- Death count for Scotland from https://www.gov.scot/coronavirus-covid-19/ (accessed 1200 `r format(today, "%d/%m/%Y")`).  
- Population counts from the Office of National Statistics (mid-year 2018).
    - UK:  https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatesforukenglandandwalesscotlandandnorthernireland, Mid-2018, spreadsheet 'MYE2-all' (accessed 1140 26/03/20) 
    - Scotland Health Board Areas:   https://statistics.gov.scot/atlas/resource?uri=http://statistics.gov.scot/id/statistical-geography/S92000003 (accessed 1200 11/03/20). 

# Doubling time calculations:

Calculated over prior 7 days using method described by *E. Vynnycky & R. White (2010) An Introduction to Infectious Disease Modelling*, page 74.  
  
Confidence intervals calculated using bootstrapping of a simulated dataset with Poisson error structure, using method published here: https://doi.org/10.1101/2020.02.05.20020750.  


# Caveats
* Case count data are affected by any changes in testing strategy or testing effort over time and/or any variation in testing strategy or testing effort between regions.  
* Case count data are likely a substantial under-representation of the true number of COVID-19 infections.  
* Death data are considered more reliable but may lag behind case data by as much as 3 weeks.  
* However, death data for London and rUKxL cannot be disaggregated. Nor can death data for Scottish Health Boards. Therefore more detailed analyses using death data are not currently possible.  
